---
title: "Simulation study parametric"
author: "Alexandre Delporte"
date: "2025-02-12"
output: html_document
---


Estimation with parametric omega function

# Set up for simulation study
```{r, cache=TRUE}
library(here)
library(htmlwidgets)
source(file.path(here("R","simulation_study","local","set_up.R")))  
```

# Low frequency high error

First, we will fix the new parameters that are introduced in the model, and estimate the others, that is $\tau$ and $\nu$.

```{r}


data_lf_he_prep=data_list[["data_lf_he"]]

H=array(rep(SIGMA_OBS_HIGH^2*diag(2),length(data_lf_he_prep$time)),dim=c(2,2,length(data_lf_he_prep$time)))
# Estimate from simulated data ------------------

formulas <- list(tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),a=~1,b=~1,D0=~1,D1=~1,sigma_D=~1,sigma_theta=~1)



colnames(data_lf_he_prep)[colnames(data_lf_he_prep)==c("observed_BoundaryDistance","observed_BoundaryAngle")]<-
  c("BoundaryDistance","BoundaryAngle")

crcvm_cubic_lf_he<- SDE$new(formulas = formulas,
                            data = data_lf_he_prep,type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,A,B,D0,D1,SIGMA_D,SIGMA_THETA),
                    fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data = list("H"=H))


#fit
crcvm_cubic_lf_he$fit(method="BFGS")
```
```{r}

crcvm_cubic_lf_he$get_all_plots(show_CI="simultaneous")
exp(as.list(crcvm_cubic_lf_he$tmb_rep(),what="Estimate")$log_sigma_obs)

```

Now we use interpolated distances and angles at high frequency to approximate better the likelihood.
First we compute and plot the interpolation.

```{r}
#prepare data for interpolation
data_lf_he_prep=data_list[["data_lf_he"]][,c("ID","y1","y2","time","observed_BoundaryDistance","observed_BoundaryAngle")]

colnames(data_lf_he_prep)[colnames(data_lf_he_prep) %in% c("observed_BoundaryDistance","observed_BoundaryAngle")]<-
  c("BoundaryDistance","BoundaryAngle")
n_step=5

#compute spline interpolation
interpolation_data<-interpolate_BoundaryMetrics(data_lf_he_prep,response=c("y1","y2"),
                                                border,n_step,n_cores=6,df_ratio=0.95)


#create dataframe with interpolated values
interpolated_x_vec=as.vector(t(interpolation_data$y1))
interpolated_y_vec=as.vector(t(interpolation_data$y2))
interpolated_distance_vec=as.vector(t(interpolation_data$BoundaryDistance))
interpolated_angle_vec=as.vector(t(interpolation_data$BoundaryAngle))
interpolation_time=as.vector(t(interpolation_data$time))
ID=unlist(lapply(unique(data_lf_he_prep$ID), function(id) {
  n=nrow(data_lf_he_prep[data_lf_he_prep$ID==id,])
  rep(id,n*n_step)
}))

#combine dataframes
interpolated_df=data.frame("time"=interpolation_time,"ID"=ID,"y1"=interpolated_x_vec,
                           "y2"=interpolated_y_vec,
                           "BoundaryDistance"=interpolated_distance_vec,
                           "BoundaryAngle"=interpolated_angle_vec)
interpolated_df$type="Interpolated"
data_lf_he_prep$type="Observed"
true_data_lf_he=data_sub[,c("ID","y1","y2","time","true_BoundaryDistance","true_BoundaryAngle")]
colnames(true_data_lf_he)[colnames(true_data_lf_he) %in% c("true_BoundaryDistance","true_BoundaryAngle")]<-
  c("BoundaryDistance","BoundaryAngle")
true_data_lf_he$type="True"

combined_data<-rbind(interpolated_df,data_lf_he_prep,true_data_lf_he)


# Create plot with legends for shape and linetype
trajectory_plot = ggplot() +geom_sf(data=border$geometry,fill="grey")+
  geom_path(data = combined_data[combined_data$time < 11, ], 
             aes(y1,y2, col = ID, linetype = type), size = 1) +
  
  # Customize the legend
  labs(color = "Trajectory ID", 
       shape = "Point Type", 
       linetype = "Line Type", 
       x = "Time", y = "Boundary Distance", 
       title = "Interpolated vs. Observed trajectories") +
  scale_shape_manual(values = c(1:3))+
  theme_minimal() +
  theme(legend.position = "right")  # Adjust legend position



distance_plot = ggplot() +
  # Scatter plot for observed BoundaryDistance (Shape Legend)
  geom_line(data = combined_data[combined_data$time < 11, ], 
             aes(time, BoundaryDistance, col = ID, linetype = type), size = 1) +
  
  # Customize the legend
  labs(color = "Trajectory ID", 
       shape = "Point Type", 
       linetype = "Line Type", 
       x = "Time", y = "Boundary Distance", 
       title = "Interpolated vs. True Boundary Distance") +
  
  theme_minimal() +
  theme(legend.position = "right")  # Adjust legend position


# Create plot with legends for shape and linetype
angle_plot = ggplot() +
  # Scatter plot for observed BoundaryDistance (Shape Legend)
  geom_line(data = combined_data[combined_data$time < 11 & combined_data$ID==1, ], 
             aes(time, BoundaryAngle, col = ID, linetype = type), size = 1) +
  
  # Customize the legend
  labs(color = "Trajectory ID", 
       shape = "Point Type", 
       linetype = "Line Type", 
       x = "Time", y = "Boundary Angle", 
       title = "Interpolated vs. True Boundary Angle") +
  
  theme_minimal() +
  theme(legend.position = "right")  # Adjust legend position

```

```{r}
# Display the plot
dyn_trajectory_plot=ggplotly(trajectory_plot)

dyn_distance_plot=ggplotly(distance_plot)
dyn_angle_plot=ggplotly(angle_plot)

saveWidget(dyn_trajectory_plot,"interpolated_trajectory.html")
saveWidget(dyn_distance_plot,"interpolated_distance.html")
saveWidget(dyn_angle_plot,"interpolated_angle.html")
```

```{r}

H=array(rep(SIGMA_OBS_HIGH^2*diag(2),length(data_lf_he_prep$time)),dim=c(2,2,length(data_lf_he_prep$time)))

#interpolated_distance[interpolated_distance==0] <- 0.05
crcvm_cubic_lf_he<- SDE$new(formulas = formulas,
                            data = data_lf_he_prep,type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,A,B,D0,D1,SIGMA_D,SIGMA_THETA),
                    fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data = list("H"=H,"interpolated_distance"=interpolation_data$BoundaryDistance,
                                              "interpolated_angle"=interpolation_data$BoundaryAngle))
crcvm_cubic_lf_he$update_lambda(c(1/0.5^2,1/0.5^2))


#fit
crcvm_cubic_lf_he$fit(method="BFGS")
```

```{r}

crcvm_cubic_lf_he$get_all_plots(show_CI="simultaneous")

```
```{r}

get_estimates<- function(sde,include_sigma_obs=FALSE) {
  
  post_coeff=sde$post_coeff(n_post=10000)
  post_par=list(
  "tau"=exp(post_coeff$coeff_fe[,"tau.(Intercept)"]),
  "nu"=exp(post_coeff$coeff_fe[,"nu.(Intercept)"]),
  "sigma_tau"=1/sqrt(exp(post_coeff$log_lambda[,"tau.s(ID)"])),
  "sigma_nu"=1/sqrt(exp(post_coeff$log_lambda[,"nu.s(ID)"])))
  
  if (include_sigma_obs) {
    post_par$sigma_obs=1000*exp(post_coeff$log_sigma_obs)
  }

  mean_par=lapply(post_par,mean)
  sd_par=lapply(post_par,sd)
  quant_par <- lapply(post_par,
                      quantile, probs = c(0.025, 0.975))
  
  return(list("estimates"=mean_par,"std.error"=sd_par,"quantiles"=quant_par))
  
}

get_estimates(crcvm_cubic_lf_he)

```

Now we try to estimate the parametric function for omega.

```{r}
H=array(rep(SIGMA_OBS_HIGH^2*diag(2),length(data_lf_he_prep$time)),dim=c(2,2,length(data_lf_he_prep$time)))

crcvm_cubic_lf_he<- SDE$new(formulas = formulas,
                            data = data_lf_he_prep,type = "CRCVM_SSM",
                    response = c("y1","y2"),fixpar=c("sigma_D","sigma_theta"),
                    par0 = c(1,1,1,1,2,1.5,SIGMA_D,SIGMA_THETA),
                    other_data = list("H"=H,"interpolated_distance"=interpolation_data$BoundaryDistance,
                                              "interpolated_angle"=interpolation_data$BoundaryAngle))


#fit
crcvm_cubic_lf_he$fit(method="BFGS")
```

```{r}

crcvm_cubic_lf_he$get_all_plots(show_CI="simultaneous")

```

# Low frequency low error 

```{r}
data_lf_le_prep=data_list[["data_lf_le"]]

colnames(data_lf_le_prep)[colnames(data_lf_le_prep)==c("observed_BoundaryDistance","observed_BoundaryAngle")]<-
  c("BoundaryDistance","BoundaryAngle")


H=array(rep(SIGMA_OBS_LOW^2*diag(2),length(data_lf_le$time)),dim=c(2,2,length(data_lf_le$time)))
# Estimate from simulated data ------------------

formulas <- list(tau=~1,
                 nu=~1,a=~1,b=~1,D0=~1,D1=~1,sigma_D=~1,sigma_theta=~1)

crcvm_cubic_lf_le<- SDE$new(formulas = formulas,data = data_lf_le_prep,
                              type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,A,B,D0,D1,SIGMA_D,SIGMA_THETA),fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("H"=H))

crcvm_cubic_lf_le$fit(method="BFGS")
```

```{r}

crcvm_cubic_lf_le$get_all_plots(show_CI="simultaneous")

```

```{r}
#compute spline interpolation
interpolation_data<-interpolate_BoundaryMetrics(data_lf_le_prep,response=c("y1","y2"),
                                                border,n_step=1,n_cores=6,k=100)


H=array(rep(SIGMA_OBS_LOW^2*diag(2),length(data_lf_le_prep$time)),dim=c(2,2,length(data_lf_le_prep$time)))

crcvm_cubic_lf_le<- SDE$new(formulas = formulas,
                            data = data_lf_le_prep,type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,A,B,D0,D1,SIGMA_D,SIGMA_THETA),
                    fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data = list("H"=H,"interpolated_distance"=interpolation_data$BoundaryDistance,
                                              "interpolated_angle"=interpolation_data$BoundaryAngle))


#fit
crcvm_cubic_lf_le$fit(method="BFGS")
```

```{r}

crcvm_cubic_lf_le$get_all_plots(show_CI="simultaneous")

```

Let's now try with high frequency data.

# High frequency high error

```{r}
H=array(rep(SIGMA_OBS_HIGH^2*diag(2),length(data_hf_he$time)),dim=c(2,2,length(data_hf_he$time)))

formulas <- list(tau=~1,
                 nu=~1,a=~1,b=~1,D0=~1,D1=~1,sigma_D=~1,sigma_theta=~1)


crcvm_cubic_hf_he_fixed<- SDE$new(formulas = formulas,data = data_hf_he[data_hf_he$DistanceShore>0.01 & data_hf_he$ID %in% c(1:6),],type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 =c(1,1,A,B,D0,D1,SIGMA_D,SIGMA_THETA),
                    fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("H"=H))

crcvm_cubic_hf_he_fixed$fit(method="BFGS")
```

```{r}
crcvm_cubic_hf_he_fixed$get_all_plots()
exp(as.list(crcvm_cubic_hf_he_fixed$tmb_rep(),what="Estimate")$log_sigma_obs)
```

Now we estimate the parameters for the function omega with the true distances and angles.

```{r}

data_hf_he_prep=data_list[["data_hf_he"]]


colnames(data_hf_he_prep)[colnames(data_hf_he_prep)==c("true_BoundaryDistance","true_BoundaryAngle")]<-
  c("BoundaryDistance","BoundaryAngle")

H=array(rep(SIGMA_OBS_HIGH^2*diag(2),length(data_hf_he_prep$time)),dim=c(2,2,length(data_hf_he_prep$time)))

formulas <- list(tau=~1,
                 nu=~1,a=~1,b=~1,D0=~1,D1=~1,sigma_D=~1,sigma_theta=~1)


crcvm_cubic_hf_he<- SDE$new(formulas = formulas,data = data_hf_he_prep,
                            type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,1,1,1,1,1,0.2),
                    other_data=list("H"=H))

#fit
crcvm_cubic_hf_he$fit(method="BFGS")
```


```{r}
crcvm_cubic_hf_he$get_all_plots()
exp(as.list(crcvm_cubic_hf_he$tmb_rep(),what="Estimate")$log_sigma_obs)
```


# High frequency low error

First fix the parametric function and estimate the other parameters of the SDE.
```{r}

H=array(rep(SIGMA_OBS_LOW^2*diag(2),length(data_hf_le$time)),dim=c(2,2,length(data_hf_le$time)))

formulas <- list(tau=~1,
                 nu=~1,a=~1,b=~1,D0=~1,D1=~1,sigma_D=~1,sigma_theta=~1)


crcvm_cubic_hf_le_fixed<- SDE$new(formulas = formulas,data = data_hf_le[data_hf_le$ID %in% c(1:6),],
                            type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,A,B,D0,D1,SIGMA_D,SIGMA_THETA),
                    fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("H"=H))

#fit
crcvm_cubic_hf_le_fixed$fit(method="BFGS")
```

```{r}
crcvm_cubic_hf_le_fixed$get_all_plots()
exp(as.list(crcvm_cubic_hf_le_fixed$tmb_rep(),what="Estimate")$log_sigma_obs)
```



Now we try to estimate D0 and a in the parametric function.

```{r}

data_hf_le_prep=data_list[["data_hf_le"]]


colnames(data_hf_le_prep)[colnames(data_hf_le_prep)==c("true_BoundaryDistance","true_BoundaryAngle")]<-
  c("BoundaryDistance","BoundaryAngle")

H=array(rep(SIGMA_OBS_LOW^2*diag(2),length(data_hf_le$time)),dim=c(2,2,length(data_hf_le$time)))

formulas <- list(tau=~1,
                 nu=~1,a=~1,b=~1,D0=~1,D1=~1,sigma_D=~1,sigma_theta=~1)


crcvm_cubic_hf_le<- SDE$new(formulas = formulas,data = data_hf_le_prep,
                            type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,1,1,1,2,1,0.2),
                    other_data=list("H"=H))

#fit
crcvm_cubic_hf_le$fit(method="BFGS")
```


```{r}
crcvm_cubic_hf_le$get_all_plots()
exp(as.list(crcvm_cubic_hf_le$tmb_rep(),what="Estimate")$log_sigma_obs)
```


```{r}
ggplot()+geom_point(data=data_hf_le,aes(time,theta),col="blue",alpha=0.5,size=0.1)+
  geom_point(data=data,aes(time,theta),col="red",alpha=0.2,size=0.1)+facet_wrap(~ID)

```
```{r}
data_hf_le$omega=fomega(data_hf_le[,c("theta","DistanceShore")])

ggplot()+geom_point(data=data_hf_le,aes(time,omega),col="blue",alpha=0.5,size=0.1)+
  geom_point(data=data,aes(time,omega),col="red",alpha=0.2,size=0.1)+facet_wrap(~ID)
```



